Natural selection in the Health and Retirement Study

### 2023-09-19 hrs-selection.qmd
Return to RLRS as a dep var

---
title: "Natural selection in the Health and Retirement Study"
author: "David Hugh-Jones"
abstract: |
  I investigate natural selection on polygenic scores
  in the contemporary US, using the Health and Retirement Study. Results
  partially support the economic theory of fertility as an explanation for
  natural selection: among both white and black respondents,
  scores which correlate negatively (positively) with education are
  selected for (against). However, unlike in the UK, there is little evidence
  that selection effects are larger among low-income or less educated people.
date: "September 2023"
format: 
  pdf:
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{placeins}
    geometry:
    - top=2in
    - left=1.25in
    - right=1.25in
    - heightrounded
editor: visual
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk: 
    echo: false
bibliography: bibliography-hrs.bib
---

@@

rand <- rand |> 
  select(ethnicity, all_of(pgs), all_of(pcs), weight, raevbrn, raedyrs,
         r10sibs, edu, income_med, married, agefb, rabyear,
         raehsamp, raestrat, parent_weight)

@@

Here, I test the same theory in the US population, using the Health and Retirement Survey [@hrsrandfam2023; @hrsrand2023]. This gives an independent test of the theory. The HRS is more representative of the population than UK Biobank, which addresses one potential weakness of the previous paper. Also, the HRS provides precalculated polygenic scores [@hrspgs2020] for both black and (non-hispanic) white participants, so I can check whether the human capital theory of natural selection works in both these ethnicities.

Economic theory treats all humans as the same. Once you have controlled appropriately for people's preferences, resources and beliefs, there should be no difference between ethnicities. So, the economic theory of fertility predicts that patterns of natural selection should be the same in both white and black ethnic groups. Of course, ethnic groups may have different economic characteristics. Also, most of the polygenic scores were created using people of European ancestry. This means they are less predictive for people with African ancestry. If natural selection works via scores' correlation with phenotypes, then we would expect these correlations to be smaller among black people.

@@

PGS are rescaled to zero mean and unit variance within each ethnic group, so effect sizes are not directly comparable between ethnic groups, but are measured in standard deviations of the within-ethnic-group score. In all regressions using PGS, I control for ten principal components of the DNA array data.

I use realised fertility (number of children ever born) as the key dependent variable. I use this rather than relative lifetime reproductive success for ease of interpretation and because I don't have data for dates when children were born. My focus is on examining patterns of natural selection across PGS, rather than estimating their effects on PGS over time; raw fertility is good enough for this.

@@

  form_pgs <- map(pgs, \(x) reformulate(c(x, pcs), response = "raevbrn"))
  mod_pgs[[eth]] <- map(form_pgs, 
                        \(x) svyglm_quiet(x, r_eth)
                        )

@@

  # the colon interaction estimates the effect separately within each group
  form_pgs_x_ed <- paste0(pgs, ":edu") |>
    map(\(x) reformulate(c(x, "edu", pcs), response = "raevbrn"))
  mod_pgs_x_ed[[eth]]  <-
    map(form_pgs_x_ed, 
        \(x) svyglm_quiet(x, r_eth)
        )

@@

  form_pgs_x_inc <- paste0(pgs, ":income_med") |>
    map(\(x) reformulate(c(x, "income_med", pcs), response = "raevbrn"))
  mod_pgs_x_inc[[eth]]  <-
    map(form_pgs_x_inc, 
        \(x) svyglm_quiet(x, r_eth)
        )

@@

  form_pgs_x_mar <- paste0(pgs, ":married") |>
    map(\(x) reformulate(c(x, "married", pcs), response = "raevbrn"))
  mod_pgs_x_mar[[eth]]  <-
    map(form_pgs_x_mar, 
        \(x) svyglm_quiet(x, r_eth)
        )

@@

  form_pgs_x_agefb <- paste0(pgs, ":agefb") |>
    map(\(x) reformulate(c(x, "agefb", pcs), response = "raevbrn"))
  mod_pgs_x_agefb[[eth]]  <-
    map(form_pgs_x_agefb, 
        \(x) svyglm_quiet(x, r_eth)
        )

@@

  cor_fert_edu_phen[[eth]] <- 
    svyglm_quiet(raevbrn ~ raedyrs, r_eth) |> 
    coef() |> 
    pluck("raedyrs")
}

@@

n_tests <- nrow(tidy_pgs_eth)
n_sig_bonf <- sum(tidy_pgs_eth$p.value < 0.05/n_tests)
if (n_sig_bonf != 0) stop("More/less than 0 score was Bonferroni significant")

@@

calc_eth_cor <- function (weights, data) {
  coef_pgs_white <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raevbrn")) |> 
        map(\(x) lm(x, data, weights = weights, 
                    subset = ethnicity == "white")) |> 
        map2_dbl(pgs, \(x, y) coef(x)[[y]])
  coef_pgs_black <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raevbrn")) |> 
        map(\(x) lm(x, data, weights = weights, 
                    subset = ethnicity == "black")) |> 
        map2_dbl(pgs, \(x, y) coef(x)[[y]])

@@

I estimated effect sizes of PGS correlations with realised fertility among black and white respondents separately.[^1] Appendix @fig-fertility shows effect sizes for white respondents only; power is too low for individual PGS estimates to be informative in the black sample. Standard errors are large because of the relatively low sample sizes, and no scores are significant at Bonferroni-corrected p \< 0.05/`r n_tests`. But I am most concerned with looking at patterns across scores rather than judging the significance of individual scores.

[^1]: Throughout, I use "effect" as shorthand for "regression coefficient". No claim about causality is implied. Recall that natural selection involves correlation, not necessarily causation, between selected characteristics and fertility.

@@

  cors_fert_edu[[eth]] <- 
    withReplicates(rand_boot, theta = calc_cors_edu, var = "raevbrn")
  cors_sibs_edu[[eth]] <- 
    withReplicates(r_parents_boot, theta = calc_cors_edu, var = "r10sibs")
}

@@

calc_eth_diff <- function (weights, data) {
  coef_pgs_white <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raevbrn")) |> 
        map(\(x) lm(x, data, weights = weights, 
                    subset = ethnicity == "white")) |> 
        map_dbl(\(x) coef(x)[[2]])
  coef_pgs_black <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raevbrn")) |> 
        map(\(x) lm(x, data, weights = weights, 
                    subset = ethnicity == "black")) |> 
        map_dbl(\(x) coef(x)[[2]])

@@

@fig-fertility-edu plots each PGS's partial correlation with fertility against its partial correlation with educational attainment. The relationship is negative among whites (correlation `r cors_fert_edu$white[1]`, bootstrapped 95% C.I. `r ci_boot_fert$white[1]` to `r ci_boot_fert$white[2]`) and negative but insignificant among blacks (correlation `r cors_fert_edu$black[1]`, bootstrapped 95% C.I. `r ci_boot_fert$black[1]` to `r ci_boot_fert$black[2]`). Bootstraps are used so as to make inferences from the sample of respondents (the different PGS are not a "sample" of anything).

```{r}
#| label: fig-fertility-edu
#| fig-cap: |
#|   Scatterplot of PGS effect sizes on fertility and years of
#|   education. Controls include 10 principal components of genetic array
#|   data. Dashed lines show linear regressions.
#| fig-subcap:
#|   - ""
#|   - ""
#| layout-nrow: 2
#| column: page

for (eth in eths) {
  ggp <- ggplot(tidy_pgs_joined[[eth]], aes(estimate.edyrs, estimate.fert)) +
          geom_hline(yintercept = 0, color = "grey80") +
          geom_vline(xintercept = 0, color = "grey80") +
          geom_point() +
          geom_smooth(method = "lm", formula = y ~ x, se = FALSE, 
                      linetype = "dashed", colour = "red", linewidth = 0.5) +
          scale_y_continuous(breaks = seq(-0.35, 0.15, 0.05),
                             labels = scales::label_number(0.01)) +
          coord_cartesian(xlim = c(-.3, .75)) +
          theme_light() + 
          theme(panel.grid = element_blank()) +
          labs(
            title = titles[eth],
            x = "Effect on years of education",
            y = "Effect on realized fertility"
          )
  print(ggp)
}

@@

  forms <- paste0(pgs, ":", term) |>
    map(\(x) reformulate(c(x, term, pcs), response = "raevbrn"))
  mods <- map(forms, \(x) lm(x, data, weights = w))
  cfs <- map(mods, coef)
  cfs_low <- map_dbl(cfs, \(x) x[str_detect(names(x), low_group)])
  cfs_high <- map_dbl(cfs, \(x) x[str_detect(names(x), high_group)])

@@

: Bootstrap estimates and 95% confidence intervals for mean absolute effect sizes of PGS on fertility in "low" and "high" groups among white respondents. Groups are: 0-12 years education vs. 13-17 years; below vs. above median income; all others vs. married; below vs. above median age at first birth. `r n_reps_x` bootstraps. {#tbl-x-bootstraps tbl-colwidths="[16,28,28,28]"}

The economic theory of fertility in @hugh2022human implies that correlations between education and fertility should be more negative for lower-income/education people, single parents and people who have children earlier.[^2] This in turn implies that correlations between PGS effects on education and on fertility should be more negative among these groups. I tested this with bootstraps, but confidence intervals were always too wide to be informative.

@@

The results here provide qualified support for the economic theory of fertility as an explanation for contemporary natural selection in humans. PGS which predict less education are being selected for, and PGS which predict more education are being selected against. PGS effects on fertility also appear larger for low-income groups and unmarried respondents. However, there is little evidence for larger effects among people with lower education or younger parents. This may be due to the low sample size. But in the UK, the between-group differences were large [@hugh2022human]; differences that big would surely have been visible here. Lastly, the smaller black sample makes most tests inconclusive for this population: I can only say that the data do not reject a negative association between PGS correlations with fertility and on education.

@@

```{r}
#| label: fig-fertility
#| fig-cap: |
#|  Effects of PGS on realized fertility among white respondents. 
#|  Values are partial correlations of a one standard deviation 
#|  (within ethnicity) change of the PGS with fertility, controlling 
#|  for 10 principal components of genomic array data. 
#| fig-height: 6

tidy_pgs$white |>
  mutate(
    term = nice_names[term],
    term = fct_reorder(term, estimate)
  ) |>
  ggplot(aes(estimate, term)) +
    geom_vline(xintercept = 0) +
    geom_pointrange(aes(xmin = conf.low, xmax = conf.high), size = 0.3) +
    theme_light() + 
    labs(
      x = "", y = "",
      title = "Effects of polygenic scores on fertility"
    ) +
    theme(
      axis.text.y = element_text(size = 7),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      legend.position = "none"
    )
```

@@

```{r}
#| label: fig-x-edu
#| fig-cap: |
#|   Boxplots of polygenic score effects on fertility by education
#|   among white respondents. Effect sizes are
#|   absolute sizes of effects of a one standard deviation (within-ethnicity) 
#|   change of the PGS on fertility, controlling for 10 principal 
#|   components of genetic array data, estimated within the low/high
#|   education group. Boxes show quartiles.

tidy_pgs_x_ed$white |> 
  mutate(
    Education = str_remove(edu, "edu")
  ) |> 
  ggplot(aes(Education, abs(estimate), color = Education)) +
    geom_x_boxplot + geom_x_point +
    labs(
      x = "Years of education",
      y = "Absolute effect size",
      title = paste("PGS effects on fertility by education", 
                    calc_n(mod_pgs_x_ed))
    ) +
    themes_x
```

```{r}
#| label: fig-x-inc
#| fig-cap: |
#|   Boxplots of polygenic score effects on fertility by income
#|   among white respondents. Effect sizes are
#|   absolute sizes of effects of a one standard deviation (within-ethnicity) 
#|   change of the PGS on fertility, controlling for 10 principal 
#|   components of genetic array data, estimated within the low/high income
#|   group. Boxes show quartiles.

tidy_pgs_x_inc$white |> 
  mutate(
    Income = fct_relevel(income_med, c("Low", "High"))
  ) |> 
  ggplot(aes(Income, abs(estimate), color = Income)) +
    geom_x_boxplot + geom_x_point +
    labs(
      x = "Income",
      y = "Absolute effect size",
      title = paste("PGS effects on fertility by income", 
                    calc_n(mod_pgs_x_inc))
    ) +
    themes_x
```

```{r}
#| label: fig-x-mar
#| fig-cap: |
#|   Boxplots of polygenic score effects on fertility by marital status
#|   among white respondents. Effect sizes are
#|   absolute sizes of effects of a one standard deviation (within-ethnicity) 
#|   change of the PGS on fertility, controlling for 10 principal 
#|   components of genetic array data, estimated within the low/high income
#|   group. Boxes show quartiles.

tidy_pgs_x_mar$white |> 
    mutate(
      married = fct_relevel(married, "Other", "Married")
    ) |> 
    ggplot(aes(married, abs(estimate), color = married)) +
      geom_x_boxplot + geom_x_point +
      labs(
          x = "Married",
          y = "Absolute effect size",
          title = paste("PGS effects on fertility by marital status", 
                        calc_n(mod_pgs_x_mar))
      ) +
      themes_x

@@

```{r}
#| label: fig-x-agefb
#| fig-cap: |
#|   Boxplots of polygenic score effects on fertility by age at 
#|   first birth (median split) among white respondents. Effect sizes are
#|   absolute sizes of effects of a one standard deviation (within-ethnicity) 
#|   change of the PGS on fertility, controlling for 10 principal 
#|   components of genetic array data, estimated within the low/high income
#|   group. Boxes show quartiles.

tidy_pgs_x_agefb$white |> 
    mutate(
      agefb = fct_relevel(agefb, "Low", "High")
    ) |> 
    ggplot(aes(agefb, abs(estimate), color = agefb)) +
      geom_x_boxplot + geom_x_point +
      labs(
          x = "Age at first birth",
          y = "Absolute effect size",
          title = paste("PGS effects on fertility by age at first birth", 
                        calc_n(mod_pgs_x_agefb))
      ) +
      themes_x

>>>>

---
title: "Natural selection in the Health and Retirement Study"
author: "David Hugh-Jones"
abstract: |
  I investigate natural selection on polygenic scores
  in the contemporary US, using the Health and Retirement Study. Results
  partially support the economic theory of fertility as an explanation for
  natural selection: among both white and black respondents,
  scores which correlate negatively (positively) with education are
  selected for (against). Selection effects are larger among low-income
  and unmarried parents, but not among younger parents or those with less 
  education.
date: "September 2023"
format: 
  pdf:
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{placeins}
    geometry:
    - top=2in
    - left=1.25in
    - right=1.25in
    - heightrounded
editor: visual
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk: 
    echo: false
bibliography: bibliography-hrs.bib
---

@@

rand <- rand |> 
  select(ethnicity, all_of(pgs), all_of(pcs), weight, raevbrn, rlrs,
         raedyrs, r10sibs, edu, income_med, married, agefb, rabyear,
         raehsamp, raestrat, parent_weight)

@@

Here, I make an independent test of the theory in the US population, using the Health and Retirement Survey [@hrsrandfam2023; @hrsrand2023]. The HRS is more representative of the population than UK Biobank, which addresses one potential weakness of the previous paper. The HRS also provides precalculated polygenic scores [@hrspgs2020] for both black and (non-hispanic) white participants, so I can check whether the human capital theory predicts patterns of natural selection in both these ethnicities.

@@

PGS are rescaled to zero mean and unit variance within each ethnic group, so effect sizes are not directly comparable between ethnic groups, but are measured in standard deviations of the within-ethnic-group score. In all regressions using PGS, I control for ten within-ethnicity principal components of the DNA array data.

The key dependent variable is relative lifetime reproductive success (RLRS): number of children ever born, divided by mean number of children of people born in the same year. RLRS is calculated pooling ethnicities, i.e. treating them as members of the same biological population.

@@

  form_pgs <- map(pgs, \(x) reformulate(c(x, pcs), response = "rlrs"))
  mod_pgs[[eth]] <- map(form_pgs, 
                        \(x) svyglm_quiet(x, r_eth)
                        )

@@

  # the colon interaction estimates the effect separately within each group
  form_pgs_x_ed <- paste0(pgs, ":edu") |>
    map(\(x) reformulate(c(x, "edu", pcs), response = "rlrs"))
  mod_pgs_x_ed[[eth]]  <-
    map(form_pgs_x_ed, 
        \(x) svyglm_quiet(x, r_eth)
        )

@@

  form_pgs_x_inc <- paste0(pgs, ":income_med") |>
    map(\(x) reformulate(c(x, "income_med", pcs), response = "rlrs"))
  mod_pgs_x_inc[[eth]]  <-
    map(form_pgs_x_inc, 
        \(x) svyglm_quiet(x, r_eth)
        )

@@

  form_pgs_x_mar <- paste0(pgs, ":married") |>
    map(\(x) reformulate(c(x, "married", pcs), response = "rlrs"))
  mod_pgs_x_mar[[eth]]  <-
    map(form_pgs_x_mar, 
        \(x) svyglm_quiet(x, r_eth)
        )

@@

  form_pgs_x_agefb <- paste0(pgs, ":agefb") |>
    map(\(x) reformulate(c(x, "agefb", pcs), response = "rlrs"))
  mod_pgs_x_agefb[[eth]]  <-
    map(form_pgs_x_agefb, 
        \(x) svyglm_quiet(x, r_eth)
        )

@@

  cor_fert_edu_phen[[eth]] <- 
    svyglm_quiet(rlrs ~ raedyrs, r_eth) |> 
    coef() |> 
    pluck("raedyrs")
}

@@

n_tests <- nrow(tidy_pgs_eth)
n_sig_bonf <- sum(tidy_pgs_eth$p.value < 0.05/n_tests)

@@

calc_eth_cor <- function (weights, data) {
  coef_pgs_white <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "rlrs")) |> 
        map(\(x) lm(x, data, weights = weights, 
                    subset = ethnicity == "white")) |> 
        map2_dbl(pgs, \(x, y) coef(x)[[y]])
  coef_pgs_black <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "rlrs")) |> 
        map(\(x) lm(x, data, weights = weights, 
                    subset = ethnicity == "black")) |> 
        map2_dbl(pgs, \(x, y) coef(x)[[y]])

@@

I estimated effect sizes of PGS correlations with RLRS among black and white respondents separately.[^1] Appendix @fig-rlrs shows effect sizes for white respondents only; power is too low for individual PGS estimates to be informative in the black sample. Standard errors are large because of the relatively low sample sizes, and only `r n_sig_bonf` score is significant at Bonferroni-corrected p \< 0.05/`r n_tests`. But I am most concerned with looking at patterns across scores rather than judging the significance of individual scores.

[^1]: Throughout, "effect" is shorthand for "regression coefficient". No claim about causality is implied. Recall that natural selection involves correlation, not necessarily causation, between selected characteristics and fertility.

@@

  cors_fert_edu[[eth]] <- 
    withReplicates(rand_boot, theta = calc_cors_edu, var = "rlrs")
  cors_sibs_edu[[eth]] <- 
    withReplicates(r_parents_boot, theta = calc_cors_edu, var = "r10sibs")
}

@@

calc_eth_diff <- function (weights, data) {
  coef_pgs_white <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "rlrs")) |> 
        map(\(x) lm(x, data, weights = weights, 
                    subset = ethnicity == "white")) |> 
        map_dbl(\(x) coef(x)[[2]])
  coef_pgs_black <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "rlrs")) |> 
        map(\(x) lm(x, data, weights = weights, 
                    subset = ethnicity == "black")) |> 
        map_dbl(\(x) coef(x)[[2]])

@@

@fig-rlrs-edu plots each PGS's partial correlation with RLRS against its partial correlation with educational attainment. The relationship is negative among whites (correlation `r cors_fert_edu$white[1]`, bootstrapped 95% C.I. `r ci_boot_fert$white[1]` to `r ci_boot_fert$white[2]`) and negative but insignificant among blacks (correlation `r cors_fert_edu$black[1]`, bootstrapped 95% C.I. `r ci_boot_fert$black[1]` to `r ci_boot_fert$black[2]`). Bootstraps are used so as to make inferences from the sample of respondents (the different PGS are not a "sample" of anything).

```{r}
#| label: fig-rlrs-edu
#| fig-cap: |
#|   Scatterplot of PGS effect sizes on RLRS and years of
#|   education. Controls include 10 principal components of genetic array
#|   data. Dashed lines show linear regressions.
#| fig-subcap:
#|   - ""
#|   - ""
#| layout-nrow: 2
#| column: page

for (eth in eths) {
  ggp <- ggplot(tidy_pgs_joined[[eth]], aes(estimate.edyrs, estimate.fert)) +
          geom_hline(yintercept = 0, color = "grey80") +
          geom_vline(xintercept = 0, color = "grey80") +
          geom_point() +
          geom_smooth(method = "lm", formula = y ~ x, se = FALSE, 
                      linetype = "dashed", colour = "red", linewidth = 0.5) +
          scale_y_continuous(breaks = seq(-0.35, 0.15, 0.05),
                             labels = scales::label_number(0.01)) +
          coord_cartesian(xlim = c(-.3, .75)) +
          theme_light() + 
          theme(panel.grid = element_blank()) +
          labs(
            title = titles[eth],
            x = "Effect on years of education",
            y = "Effect on RLRS"
          )
  print(ggp)
}

@@

  forms <- paste0(pgs, ":", term) |>
    map(\(x) reformulate(c(x, term, pcs), response = "rlrs"))
  mods <- map(forms, \(x) lm(x, data, weights = w))
  cfs <- map(mods, coef)
  cfs_low <- map_dbl(cfs, \(x) x[str_detect(names(x), low_group)])
  cfs_high <- map_dbl(cfs, \(x) x[str_detect(names(x), high_group)])

@@

: Bootstrap estimates and 95% confidence intervals for mean absolute effect sizes of PGS on RLRS in "low" and "high" groups among white respondents. Groups are: 0-12 years education vs. 13-17 years; below vs. above median income; all others vs. married; below vs. above median age at first birth. `r n_reps_x` bootstraps. {#tbl-x-bootstraps tbl-colwidths="[16,28,28,28]"}

The economic theory of fertility in @hugh2022human implies that correlations between education and RLRS should be more negative for lower-income/education people, single parents and people who have children earlier.[^2] This in turn implies that correlations between PGS effects on education and on RLRS should be more negative among these groups. I tested this with bootstraps, but confidence intervals were always too wide to be informative.

@@

The results here provide qualified support for the economic theory of fertility as an explanation for contemporary natural selection in humans. PGS which predict less education are being selected for, and PGS which predict more education are being selected against. PGS effects on RLRS also appear larger for low-income groups and unmarried respondents. However, there is little evidence for larger effects among people with lower education or younger parents. This may be due to the low sample size. But in the UK, the between-group differences were large [@hugh2022human]; differences that big would surely have been visible here. Lastly, the smaller black sample makes most tests inconclusive for this population: I can only say that the data do not reject a negative association between PGS correlations with RLRS and with education.

@@

```{r}
#| label: fig-rlrs
#| fig-cap: |
#|  Effects of PGS on RLRS among white respondents. 
#|  Values are partial correlations of a one standard deviation 
#|  (within ethnicity) change of the PGS with RLRS, controlling 
#|  for 10 principal components of genomic array data. 
#| fig-height: 6

tidy_pgs$white |>
  mutate(
    term = nice_names[term],
    term = fct_reorder(term, estimate)
  ) |>
  ggplot(aes(estimate, term)) +
    geom_vline(xintercept = 0) +
    geom_pointrange(aes(xmin = conf.low, xmax = conf.high), size = 0.3) +
    theme_light() + 
    labs(
      x = "", y = "",
      title = "Effects of polygenic scores on RLRS"
    ) +
    theme(
      axis.text.y = element_text(size = 7),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      legend.position = "none"
    )
```

@@

```{r}
#| label: fig-x-edu
#| fig-cap: |
#|   Boxplots of polygenic score effects on RLRS by education
#|   among white respondents. Effect sizes are
#|   absolute sizes of effects of a one standard deviation (within-ethnicity) 
#|   change of the PGS on RLRS, controlling for 10 principal 
#|   components of genetic array data, estimated within the low/high
#|   education group. Boxes show quartiles.

tidy_pgs_x_ed$white |> 
  mutate(
    Education = str_remove(edu, "edu")
  ) |> 
  ggplot(aes(Education, abs(estimate), color = Education)) +
    geom_x_boxplot + geom_x_point +
    labs(
      x = "Years of education",
      y = "Absolute effect size",
      title = paste("PGS effects on RLRS by education", 
                    calc_n(mod_pgs_x_ed))
    ) +
    themes_x
```

```{r}
#| label: fig-x-inc
#| fig-cap: |
#|   Boxplots of polygenic score effects on RLRS by income
#|   among white respondents. Effect sizes are
#|   absolute sizes of effects of a one standard deviation (within-ethnicity) 
#|   change of the PGS on RLRS, controlling for 10 principal 
#|   components of genetic array data, estimated within the low/high income
#|   group. Boxes show quartiles.

tidy_pgs_x_inc$white |> 
  mutate(
    Income = fct_relevel(income_med, c("Low", "High"))
  ) |> 
  ggplot(aes(Income, abs(estimate), color = Income)) +
    geom_x_boxplot + geom_x_point +
    labs(
      x = "Income",
      y = "Absolute effect size",
      title = paste("PGS effects on RLRS by income", 
                    calc_n(mod_pgs_x_inc))
    ) +
    themes_x
```

```{r}
#| label: fig-x-mar
#| fig-cap: |
#|   Boxplots of polygenic score effects on RLRS by marital status
#|   among white respondents. Effect sizes are
#|   absolute sizes of effects of a one standard deviation (within-ethnicity) 
#|   change of the PGS on RLRS, controlling for 10 principal 
#|   components of genetic array data, estimated within the low/high income
#|   group. Boxes show quartiles.

tidy_pgs_x_mar$white |> 
    mutate(
      married = fct_relevel(married, "Other", "Married")
    ) |> 
    ggplot(aes(married, abs(estimate), color = married)) +
      geom_x_boxplot + geom_x_point +
      labs(
          x = "Married",
          y = "Absolute effect size",
          title = paste("PGS effects on RLRS by marital status", 
                        calc_n(mod_pgs_x_mar))
      ) +
      themes_x

@@

```{r}
#| label: fig-x-agefb
#| fig-cap: |
#|   Boxplots of polygenic score effects on RLRS by age at 
#|   first birth (median split) among white respondents. Effect sizes are
#|   absolute sizes of effects of a one standard deviation (within-ethnicity) 
#|   change of the PGS on RLRS, controlling for 10 principal 
#|   components of genetic array data, estimated within the low/high income
#|   group. Boxes show quartiles.

tidy_pgs_x_agefb$white |> 
    mutate(
      agefb = fct_relevel(agefb, "Low", "High")
    ) |> 
    ggplot(aes(agefb, abs(estimate), color = agefb)) +
      geom_x_boxplot + geom_x_point +
      labs(
          x = "Age at first birth",
          y = "Absolute effect size",
          title = paste("PGS effects on RLRS by age at first birth", 
                        calc_n(mod_pgs_x_agefb))
      ) +
      themes_x
